---
title: "Data Manipulation"
output: github_document
---

```{r setup}
library(tidyverse)
```

## Load in the FAS litters data.

```{r}
litters_df = read_csv("./data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
```

## `select`

Choose some columns and not other

```{r}
select(litters_df, group, litter_number)
```

```{r}
select(litters_df, group, gd0_weight:gd_of_birth)
```

```{r}
select(litters_df, -litter_number)
```

Renaming columns
(select the column that I want to keep and in the mean while renaming those)
```{r}
select(litters_df, GROUP = group, LITTER_NUMBER = litter_number)
```

rename without select
```{r}
rename(litters_df, GROUP = group, LITTER_NUMBER = litter_number)
```

Helper functions

Select Helpers

```{r}
select(litters_df, starts_with("gd"))
```

```{r}
select(litters_df, litter_number, everything())
```

relocate did not work??

## `filter`

```{r}
filter(litters_df, gd0_weight < 22)
```

```{r}
filter(litters_df, gd_of_birth == 20)
```

```{r}
filter(litters_df, !(gd_of_birth == 20))
```

```{r}
filter(litters_df, gd0_weight >= 22, gd_of_birth == 20)
```

```{r}
filter(litters_df, group == "Mod 8")
```

```{r}
filter(litters_df, group %in% c("Con7", "Mod 8"))
```

## `mutate`

create a new variable and modify variable
```{r}
mutate(
  litters_df, 
  wt_gain = gd18_weight - gd0_weight,
  group = str_to_lower(group))
```


## `arrange`
get the column in order
```{r}
arrange(litters_df, pups_born_alive)
```

first order by pups, then in the pups order by gd0weight.
```{r}
arrange(litters_df, pups_born_alive, gd0_weight)
```

## `%>%`

```{r}
litters_data_raw = read_csv("./data/FAS_litters.csv")
litters_clean_name = janitor::clean_names(litters_data_raw)
litters_data_selected = select(litters_clean_name, -pups_survive)
litters_mutated = mutate(litters_data_selected, wt_gain = gd18_weight - gd0_weight)
litters_without_missing = drop_na(litters_mutated, gd0_weight)
```


USE THE PIPE OPERATOR INSTEAD

```{r}
litters_df = 
  read_csv("./data/FAS_litters.csv") %>%
  janitor::clean_names() %>% 
  select(-pups_survive) %>% 
  mutate(wt_gain = gd18_weight - gd0_weight) %>% 
  drop_na(gd0_weight)
```





